---
title: "Family of APPLY functions"
author: "Michael Wang"
date: "2017Äê1ÔÂ27ÈÕ"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown file for using apply, lapply and sapply

'apply' used for array/matrix and dataframe, with the form apply(x,margin,function,...), 
the output varies depends on functions.

```{r}
#================================================================
# summarise
# for array and martix, dplyr is not possible, which only deals with dataframe like object
#================================================================
str(Titanic)
Titanic
apply(Titanic,1,sum) #subtotal in each category of the 'Class' dimension
apply(Titanic,2,sum) #subtotal in each category of the 'Sex' dimennsion
apply(Titanic,3,sum) #subtotal in each category of the 'age' dimension
apply(Titanic,4,sum) #subtotal in each category of the 'suvival' dimension

apply(Titanic,c(3,4),sum) #group by 3rd and 4th dimensions, 4-d squeezed to 2-d data
apply(Titanic,c(1,2),sum) #group by 1st and 2nd dimensions, 4-d squeezed to 2-d data

```

lapply and sapply, with the form lapply/sapply(x,function,...)

'lapply' deals with list, vector and dataframe, and outpout list; 
'sapply' is exactly the same as 'lapply', excpt the output reducing to matrix, vector, or list

```{r}
str(iris)
lapply(iris[,-5],mean)  # always columwise operation, excluding the factor coloumn
sapply(iris[,-5],mean) # the result is a vector this time
apply(iris[,-5],2,mean) # need to include dimension '2'
lapply(iris[,-5],max) # max of each column, excluding the factor column
sapply(iris[,-5],max) # the result is a vector this time.
apply(iris[,-5],2,max) # need to include dimension '2' this time

```

tapply work with vector, columns of dataframe or list, summary data by group, equivalent to   group_by and summarise in dplyer, but dplyr can work on mutiple columns

```{r}
head(iris)
test<-tapply(iris$Sepal.Length,iris$Species,mean) #using tapply
class(test)

library(dplyr)
iris%>%
  group_by(Species)%>%
  summarise_each(funs(mean))  # done for all columns

iris%>%
  group_by(Species)%>%
  summarise_each(funs(mean),Sepal.Length,Sepal.Width)  # done for two columns

iris%>%
  group_by(Species)%>%
  summarise_each(funs(mean),Sepal.Length)  # done for ONE columns

iris%>%
  group_by(Species)%>%
  summarise(average_Sepal.Length=mean(Sepal.Length,na.rm=T))  # use 'summarise this time

```
source: 
  1. https://www.youtube.com/watch?v=jWjqLW-u3hc&t=1302s
  2. https://www.youtube.com/watch?v=f0U74ZvLfQo&t=610s
  3. https://www.youtube.com/watch?v=ejVWRKidi9M
  4. https://www.youtube.com/watch?v=HmBPDTtb6Bg

